---
title: "Comparison of Exponential Distribution against Central Limit Theorem"
author: "Sooraj Raju"
date: "May 8, 2016"
output: pdf_document
---

# 1. Overview

The objective of this project is to investigate the Exponential Distribution and compare it against Central Limit Theorem. 

Central Limit Theorem:
In probability theory, the central limit theorem (CLT) states that, given certain conditions, the arithmetic mean of a sufficiently large number of iterates of independent random variables, each with a well-defined expected value and well-defined variance, will be approximately normally distributed, regardless of the underlying distribution

The exponential distribution will be simulated in R with rexp(n, lambda) where lambda is the rate parameter. In theory the mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda.

We will set n = 40 and set lambda = 0.2 for all of the simulations and compare the resulting distribution of the smaple averages to the theory for 1000 simulations.
```{r, results='hide', message=FALSE, warning=FALSE }
library(ggplot2)
library(psych)
```
```{r}

set.seed(1000)
lambda <- 0.2 ## Set lambda as per instructions

## Distibution size
DisSize <- 40 

## number of simulations
NoOfSim <- 1000 

MeanOfDistrib <- 0
for (i in 1 : NoOfSim) 
  MeanOfDistrib <- c(MeanOfDistrib, mean(rexp(DisSize,lambda)))
hist(MeanOfDistrib,col="red",main="Distribution of averages of exponential distribution")
```



```{r}
varxp <- ((1/lambda)^2)/DisSize ## theoretical variance
varmean <- var(MeanOfDistrib) ## variance of the means
Mean_theoritical <- 1/lambda
```
```{r echo= FALSE}
print(sprintf("The theoretical mean is %f while mean of means is %f",Mean_theoritical,mean(MeanOfDistrib)))
print(sprintf("The theoretical variance is %f while variance of means %f",varxp,varmean))
```
```{r}

head(MeanOfDistrib)
summary(MeanOfDistrib)
describe(MeanOfDistrib)
plotdata <- data.frame(MeanOfDistrib)
plot1 <- ggplot(plotdata,aes(x = MeanOfDistrib))
plot1 <- plot1 +geom_histogram(aes(y=..density..), colour="red",fill="white")
plot1<-plot1+labs(title="Distribution of Means of exponential distribution ", y="Density")
plot1<-plot1 +stat_function(fun=dnorm,args=list( mean=1/lambda, sd=sqrt(varxp)),color = "blue", size = 1.0)
plot1<-plot1 +stat_function(fun=dnorm,args=list( mean=mean(MeanOfDistrib), sd=sqrt(varmean)),color = "green", size = 1.0)
print(plot1)
```

```{r cache=TRUE}

set.seed(1000)
lambda <- 0.2 ## Set lambda as per instructions

## Distibution size
DisSize <- 40 

## number of simulations
NoOfSim <- 100000

MeanOfDistrib <- 0
for (i in 1 : NoOfSim) 
  MeanOfDistrib <- c(MeanOfDistrib, mean(rexp(DisSize,lambda)))
hist(MeanOfDistrib,col="red",main="Distribution of averages of exponential distribution")
```



```{r}
varxp <- ((1/lambda)^2)/DisSize ## theoretical variance
varmean <- var(MeanOfDistrib) ## variance of the means
Mean_theoritical <- 1/lambda
```
```{r echo= FALSE}
print(sprintf("The theoretical mean is %f while mean of means is %f",Mean_theoritical,mean(MeanOfDistrib)))
print(sprintf("The theoretical variance is %f while variance of means %f",varxp,varmean))
```
```{r}

head(MeanOfDistrib)
summary(MeanOfDistrib)
describe(MeanOfDistrib)
plotdata <- data.frame(MeanOfDistrib)
plot1 <- ggplot(plotdata,aes(x = MeanOfDistrib))
plot1 <- plot1 +geom_histogram(aes(y=..density..), colour="red",fill="white")
plot1<-plot1+labs(title="Distribution of Means of exponential distribution ", y="Density")
plot1<-plot1 +stat_function(fun=dnorm,args=list( mean=1/lambda, sd=sqrt(varxp)),color = "blue", size = 1.0)
plot1<-plot1 +stat_function(fun=dnorm,args=list( mean=mean(MeanOfDistrib), sd=sqrt(varmean)),color = "green", size = 1.0)
print(plot1)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
